Water desalination using polyelectrolyte hydrogel: Gibbs ensemble modeling

### 2022-09-14 gibbs.tex
Update gibbs.tex

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Recently polyelectrolyte hydrogels have been proposed as draw agents for the forward osmosis desalination techniques. 
Indeed, polyelectrolyte hydrogels have the ability to absorb a big amount of water across an osmosis membrane as a result of their swelling pressure. 
On the other hand, the insoluble cross-linked network of the gel enables dewatering under the influence of a stimulus (thermal and/or mechanical).
Moreover, from a thermodynamic perspective, the polyelectrolyte hydrogel is already an osmotic membrane rejecting ions between external and internal solutions.
These three properties of the gels make it possible to use them for desalination and at the same time avoid the use of expensive membranes.
In this article, we present our recent theoretical study of the use of polyelectrolyte hydrogel for water desalination. 
We modeled the thermodynamic equilibrium between coexisting phases of the gel and supernate aqueous salt solution.
We have shown that the salinity of the supernate phase decreases during compression of the gel due to the release of absorbed in gel solution of lower salinity. 
Finally, we performed a set of simulations modeling the process of continuous decrease of solution salinity up to freshwater concentrations.
}

@@

We propose a cascade desalination process to lower supernatant salinity.
%it is worth to note that we are not limited with this process only.
For this process the gel must be commpressed/decompressed in an \emph{open} and \emph{closed} system.
By the \emph{open system} we mean that the gel is in thermodynamic equilibrium with huge, effectively \textbf{infinite}, reservoir of an aqueous solution. 
The gel freely exchanges ions with the reservoir, while their concentration in reservoir (and chemical potential) remains constant.
%, such that there are the same ion's chemical potential, the same temperature, and desired pressure difference between the gel and the reservoir. The number of  $\na$ and $\cl$ ions and gel volume $\Vgel$ are allowed to change.
By the \emph{closed system} we mean that the gel is in equilibrium with a \textbf{finite} reservoir of an aqueous solution. 
In this case the exchange of ions between the gel and the reservoir may affect their concentration in reservoir, whereas the  
%again have the same ions chemical potential, the same temperature, and desired pressure difference between the gel and the reservoir. 
%Although, the number of ions and volume of gel and reservoir may vary, 
the total number of ions (in gel and in reservoir) is constant.

Due to huge size of the reservoir, the compression of the gel in the \emph{open system} does not affect the reservoir salinity, $\cs =$ $\cna=$ $\ccl=$ $\text{Const}$ (\reffig{fig: open}), whereas the number of ions entrapped in the volume $\Vbox$ changes.
On the contrary, the compression of the gel in the \emph{closed system} changes the salinity in the reservoir, but the total number of $\na$ and $\cl$ ions in the gel and in the reservoir, \ie in the volume $\Vbox$ (\reffig{fig: closed}), remains constant.

%In both cases there are two coexisting systems with constant $\mu P T$ allowed to exchange particles and change volume, which is represented as Gibbs ensemble.

Unfortunately, there is no universal computationally effective technique to sample pressure difference, volume, and number of particles at the same. To equilibrate and sample pressure difference and volume we use molecular dynamics (MD) simulation, while equlibration and sampling of ions distribution between the gel and a reservoir is done via Monte-Carlo (MC) simulation. In this sence employed techniques is called MCMD.

To simulate processes in \emph{open system} and  \emph{closed system} two complementary models MC and MD were implemented.

\subsubsection*{Molecular Dynamics model}

Our MD model of the gel is a network of 16 linear polymer chains, each consisting of 30 monomer 
units. 
These polymer chains are connected to a diamond-like network by 8 crosslinking units.
This means, there are $\Ngel=16\cdot30 + 8=488$ gel monomers in the simulation box 
(see \reffig{fig:diamond}).
Each monomer unit of the network carries a negative elementary electric charge, thus 
simulating fully ionized polyacidic gel.
Except the particles in the network, the monovalent co- and counter-ions ions, 
$\cl$ and $\na$, are present in simulation box. 
The total electric charge of all the particles in the box is zero, therefore the amount 
of $\na$ ions is bigger than that of $\cl$ by the number of hydrogel units.
The entire simulation box setup employs periodic boundaries.
Coloumb interaction computed by P3M method.
This model is used in MD part of \emph{open system} and \emph{closed system} simulation.

For the MD part of \emph{closed system} simulation analougous simulation box is created without any polymer monomers to model the finite reservoir. 
The amount of $\na$ ions is equal to $\cl$.

By integration Langevin motion equation by Verlet algorithm for each particles we sample 
total energy of the system and pressure. 
Kinetic energy, bonded, non-bonded and Coloumb interaction contribute to the total energy of a system.
Pressure in the system calculated as in ref [Essmann et al., 1995, doi:10.1063/1.470117]
\todo{I do not know if it is needed, ML}

When computing MD the gel and reservoir box are essentially NVT-ensembles.

\subsubsection*{Monte Carlo model}

In the \emph{open system} system box with the gel and the monovalent co- and counter-ions ions, 
$\cl$ and $\na$ is allowed to exchange particles with infinite reservoir with constant chemical potential.

\dots

In the \emph{closed system} system two boxes one with the gel and the other with ions pairs only iare allowed to exchange particles.

@@

\subsection{}

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Recently polyelectrolyte hydrogels have been proposed as draw agents for the forward osmosis desalination techniques. 
Indeed, polyelectrolyte hydrogels have the ability to absorb a big amount of water across an osmosis membrane as a result of their swelling pressure. 
On the other hand, the insoluble cross-linked network of the gel enables dewatering under the influence of a stimulus (thermal and/or mechanical).
Moreover, from a thermodynamic perspective, the polyelectrolyte hydrogel is already an osmotic membrane rejecting ions between external and internal solutions.
These three properties of the gels make it possible to use them for desalination and at the same time avoid the use of expensive membranes.

In this article, we present our recent theoretical study of the usage of polyelectrolyte hydrogel for water desalination. 
Using coarse-grained model and the Gibbs ensemble, we modeled the thermodynamic equilibrium between coexisting phases of the gel and corresponding supernate aqueous 
salt solution.

We performed a sequence of step-by-step hydrogel compressions in \emph{open} and \emph{closed} systems.
The compression in \emph{open system} causes the decrease of salt ions inside the hydrogel,
while the compression in \emph{clesed system} causes the decrease of the supernatant solution.
We modeled this stepwise process of continuous decrease of seawater salinity up to freshwater concentrations.
We estimated the energy costs of the entire process as 38.2 J per production of 1 liter of potable water, which is comparable with reverse osmosis.
}

@@

Our model of the gel is a network of 16 linear polymer chains, each consisting of 30 monomer 
units. 
These polymer chains are connected to a diamond-like network by 8 crosslinking units.
This means, there are $\Ngel=16\cdot30 + 8=488$ gel monomers in the simulation box 
(see \reffig{fig:diamond}).
Each monomer unit of the network carries a negative elementary electric charge, thus 
simulating fully ionized polyacidic gel.
Except the particles in the network, the monovalent co- and counter-ions ions, 
$\cl$ and $\na$, are present in simulation box. 
The total electric charge of all the particles in the box is zero, therefore the amount 
of $\na$ ions is bigger than that of $\cl$ by the number of hydrogel units.
The entire simulation box setup employs periodic boundaries.

@@
